Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-06 popl24/intro.tex
checkpoint

Security in the MPC setting thus means that protocols cannot reveal
any secret information other than what is impicitly declassified by
the publicized output of the ideal functionality. The security model
also assumes that some subset of participants can be corrupted and
collude adversarially to possibly infer more secret information. The
accepted method of demonstrating protocol security in this setting is
to define a \emph{simulation} algorithm that runs in the ``ideal''
world which, given just the inputs of corrupted parties and the output
of the ideal functionality, is able to reconstruct information that
corrupted parties receive in their so-called \emph{views} of the
protocol running in the real world.  This implies that adversarial
views provide no information beyond what is provided by the ideal
output alone. Simulation is defined probabilistically since MPC
protocols typically rely on cryptographic and probabilistic methods.

@@

We distinguish between \emph{extensional} vs.~\emph{intensional}
properties and analysis of MPC protocols. By extensional, we mean
simulator security itself as well as analysis concerned with the use
and interaction of complete secure protocols. By intensional, we mean
analysis and inner workings of protocols themselves. While the
probabilistic formulations we develop could conceivably used
extensionally, our main focus is on automated and semi-automated
reasoning about intensional properties of protocols, both through
whole-program analysis and compositional properties of program
components that support security.

Previous work on analysis for the SecreC language
\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
extensional properties of MPC, in particular the specification and
enforcement of declassification bounds in programs that use MPC in
subprograms. This work is explicitly reminiscent of information flow
approaches such as delimited information release
\cite{10.1007/978-3-540-37621-7_9} and downgrading policies and
relaxed noninterference \cite{10.1145/1040305.1040319}. However, their
program logic assumes correctness of the underlying MPC protocols.
The Wys$^\star$ language \cite{wysstar}, based on Wysteria
\cite{rastogi2014wysteria}, has similar goals and includes a
trace-based semantics for reasoning about the extensional interactions
of MPC protocols. Their compiler also guarantees that underlying
multi-threaded protocols enforce the single-threaded source language
semantics which is an intensional property. All of these systems
leverage type systems to obtain automated security analyses.

In the realm of intensional analysis, previous work has shown how to
reduce proofs of simulator security to program equivalence assertions,
which supports applications known techniques in automated verification
\cite{8429300}. However, they focus on noninterference properties,
which are arguably sound but notably incomplete wrt MPC functionality
as we discuss above and in Section
\ref{section-hyperprop-ni}. Probabilistic Separation Logic
\cite{barthe2019probabilistic} develops a logical framework for
reasoning about probabilistic independence in programs.  They consider
several MPC protocols and show how to systematically deduce, e.g.,
probabilistic independence of input secrets and adversarial views that
are critical intensional properties of simulator security.  However,
their focus on independence does not address the subtleties of
dependence that arise during MPC protocol execution, and the authors
do not formulate or prove simulator security properties in
\cite{barthe2019probabilistic}. The $\lambda_{\mathrm{obliv}}$
language \cite{darais2019language} is perhaps most closely related to
our work. Here, a type system is used to automatically enforce
so-called probabilistic trace obliviousness and they show how to use
their system to implement several variants of oblivious RAM. However,
obliviousness is related to noninterference and does not address
allowable information leakage as in MPC, and their use of type
linearity disallows reuse of sampled probabilistic values which is
common in the MPC idiom, as our examples will illustrate. Similarly,
previous work on oblivious data structures \cite{10.1145/3498713},
while intended for sound integration into MPC protocols, focus on
expression and enforcement of obliviousness and do not account
for allowable information leakage.

@@

In Section \ref{section-minicat} we develop a new probabilistic programming
language $\minifed$ for defining synchronous distributed protocols
over the binary field. The syntax and semantics provides a succinct
account of \emph{views} and (synchronous) messaging between protocol
\emph{clients}. In this Section we consider small example protocols
including oblivious transfer, communications encryption with one-time
pads, and additive secret sharing. In Section \ref{section-pmf} we
formulate \emph{program distributions}, a formalism for expressing
dependencies between secrets and views, useful for expressing
intensional characteristics of protocols and precisely quantifying
allowable information leakage in terms of protocol outputs. On this
basis we then formulate a novel hyperproperty of program execution
traces, called \emph{noninterference modulo output ($\NIMO$)} that we
demonstrate implies passive simulator security. In Section
\ref{section-bruteforce} we define a brute force mechanism for
verifying passive security in $\minifed$ protocols, that is also
amenable to HPC optimizations via translation of $\minifed$ protocols
into Datalog \cnote{I think this Section can be on the chopping block
  for space savings, since it's based on truth table
  computation/Datalog semantics.}.

@@

In summary, our analysis takes a ``just do it'' approach- rather than
using, e.g., traditional security type or linear type disciplines to
prevent ``bad things'' from happening, we base our analysis on precise
quantification of intensional protocol behavior and especially
dependencies between secret inputs, public outputs, and adversarial
views of protocols. We ameliorate scalability issues associated with
fine-grained analysis by manually establishing compositional
properties inspired by probabilistic separation logic
\cite{barthe2019probabilistic}, obtaining a middle-ground, scalable
``proof assistant'' style approach enabled by MPC protocol paradigms.

>>>>

Security in the MPC setting thus means that protocols cannot reveal
any secret information other than what is implicitly declassified by
the publicized output of the ideal functionality. The security model
also assumes that some subset of participants can be corrupted and
collude adversarially to possibly infer more secret information. The
accepted method of demonstrating protocol security in this setting is
to define a \emph{simulation} algorithm that runs in the ``ideal''
world which, given just the inputs of corrupted parties and the output
of the ideal functionality, is able to reconstruct information that
corrupted parties receive in their so-called \emph{views} of the
protocol running in the real world.  This implies that adversarial
views provide no information beyond what is provided by the ideal
output alone. Simulation is defined probabilistically since MPC
protocols typically rely on cryptographic and probabilistic methods.

@@

We distinguish between \emph{extensional} vs.~\emph{intensional}
properties and analysis of MPC protocols. By extensional, we mean
simulator security itself as well as analysis concerned with the use
and interaction of complete secure protocols. By intensional, we mean
analysis and inner workings of protocols themselves. While the
probabilistic formulations we develop could conceivably be used
extensionally, our main focus is on automated and semi-automated
reasoning about intensional properties of protocols, both through
whole-program analysis and compositional properties of program
components that support security.

Previous work on analysis for the SecreC language
\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
extensional properties of MPC, in particular the specification and
enforcement of declassification bounds in programs that use MPC in
subprograms. This work is explicitly reminiscent of information flow
approaches such as delimited information release
\cite{10.1007/978-3-540-37621-7_9}, downgrading policies, and
relaxed noninterference \cite{10.1145/1040305.1040319}. However, their
program logic assumes correctness of the underlying MPC protocols.
The Wys$^\star$ language \cite{wysstar}, based on Wysteria
\cite{rastogi2014wysteria}, has similar goals and includes a
trace-based semantics for reasoning about the extensional interactions
of MPC protocols. Their compiler also guarantees that underlying
multi-threaded protocols enforce the single-threaded source language
semantics which is an intensional property. \mnote{We need a clearer statement here of the flaws with Wys-star specifically} All of these systems
leverage type systems to obtain automated security analyses.

In the realm of intensional analysis, previous work has shown how to
reduce proofs of simulator security to program equivalence assertions,
which supports the application of known techniques in automated verification
\cite{8429300}. However, they focus on noninterference properties,
which are arguably sound but notably incomplete wrt MPC functionality
as we discuss above and in Section
\ref{section-hyperprop-ni}. Probabilistic Separation Logic
\cite{barthe2019probabilistic} develops a logical framework for
reasoning about probabilistic independence in programs.  They consider
several MPC protocols and show how to systematically deduce, e.g.,
probabilistic independence of input secrets and adversarial views that
are critical intensional properties of simulator security.  However,
their focus on independence does not address the subtleties of
dependence that arise during MPC protocol execution, and the authors
do not formulate or prove simulator security properties in
\cite{barthe2019probabilistic}. The $\lambda_{\mathrm{obliv}}$
language \cite{darais2019language} is perhaps most closely related to
our work. Here, a type system is used to automatically enforce
so-called probabilistic trace obliviousness and they show how to use
their system to implement several variants of oblivious RAM. However,
obliviousness is related to noninterference and does not address
allowable information leakage as in MPC, and their use of type
linearity disallows the reuse of sampled probabilistic values which is
common in the MPC idiom, as our examples will illustrate. Similarly,
previous work on oblivious data structures \cite{10.1145/3498713},
while intended for sound integration into MPC protocols, focus on
expression and enforcement of obliviousness and do not account
for allowable information leakage.

@@

In Section \ref{section-minicat} we develop a new probabilistic programming
language $\minifed$ for defining synchronous distributed protocols
over the binary field. The syntax and semantics provides a succinct
account of \emph{views} and (synchronous) messaging between protocol
\emph{clients}. In this Section, we consider small example protocols
including oblivious transfer, communications encryption with one-time
pads, and additive secret sharing. In Section \ref{section-pmf} we
formulate \emph{program distributions}, a formalism for expressing
dependencies between secrets and views, useful for expressing
intensional characteristics of protocols and precisely quantifying
allowable information leakage in terms of protocol outputs. On this
basis we then formulate a novel hyperproperty of program execution
traces, called \emph{noninterference modulo output ($\NIMO$)} that we
demonstrate implies passive simulator security. In Section
\ref{section-bruteforce} we define a brute force mechanism for
verifying passive security in $\minifed$ protocols, that is also
amenable to HPC optimizations via translation of $\minifed$ protocols
into Datalog \cnote{I think this Section can be on the chopping block
  for space savings, since it's based on truth table
  computation/Datalog semantics.}.

@@

In summary, our analysis takes a ``just do it'' approach- rather than
using, e.g., traditional security type or linear type disciplines to
prevent ``bad things'' from happening, we base our analysis on precise
quantification of intensional protocol behavior and especially
dependencies between secret inputs, public outputs, and adversarial
views of protocols. We ameliorate scalability issues associated with
fine-grained analysis by manually establishing compositional
properties inspired by probabilistic separation logic
\cite{barthe2019probabilistic}, obtaining a middle-ground, scalable
``proof assistant style'' approach enabled by MPC circuit paradigms.
